Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-13 arxiv/conclusions.tex
polish section 3 and 5

We describe several open questions for future work.
First, our bound in Corollary \ref{cor_MTL_loss} involves an error term that scales down with $n_1$.
Tightening this error bound requires showing the limit of $\normFro{({Z^{(1)}}^{\top} Z^{(1)} + {Z^{(2)}}^{\top} Z^{(2)})^{-1} {Z^{(1)}}^{\top} Z^{(1)}}^2$ for two random matrices $Z^{(1)}, Z^{(2)}$ with isotropic covariances.
This requires studying the asymptotic distribution of the {singular values} of the non-symmetric matrix $({Z^{(1)}}^{\top} Z^{(1)})^{-1}{Z^{(2)}}^{\top} Z^{(2)} + 1$, which is still an open problem in random matrix theory. The eigenvalues distribution of this matrix have been obtained in \cite{Fmatrix}, which might be relevant. 
%but its singular values will follow a different distribution since the matrix is not symmetric.
 %might require new techniques beyond the current ones in random matrix theory .%\HZ{to add}.
Second, it would be interesting to extend our results to classification problems.
Several recent work have made remarkable progress in the (one-sample) high-dimensional logistic regression setting \cite{sur2019modern}.
An interesting question is to study hard parameter sharing in (multiple-sample) high-dimensional logistitic regression.

>>>>

We describe several open questions for future work.
First, our bound in Corollary \ref{cor_MTL_loss} involves an error term that scales down with $n_1$.
Tightening this error bound requires showing the limit of $\normFro{({Z^{(1)}}^{\top} Z^{(1)} + {Z^{(2)}}^{\top} Z^{(2)})^{-1} {Z^{(1)}}^{\top} Z^{(1)}}^2$ for two isotropic random matrices.
This requires studying the asymptotic distribution of the {singular values} of the non-symmetric matrix $({Z^{(1)}}^{\top} Z^{(1)})^{-1}{Z^{(2)}}^{\top} Z^{(2)}$, which is still an open problem in random matrix theory. The eigenvalue distribution of this matrix, which has been obtained in \citet{Fmatrix}, might be helpful towards resolving this problem.
%but its singular values will follow a different distribution since the matrix is not symmetric.
 %might require new techniques beyond the current ones in random matrix theory .%\HZ{to add}.
Second, it would be interesting to extend our results to classification problems.
Several recent work have made remarkable progress for logistic regression in the high-dimensional  setting, e.g. \cite{sur2019modern}.
An interesting question is to study logistic regression in a multiple-sample setting.

### 2020-10-13 arxiv/diff_cov.tex
polish section 3 and 5

Our key result characterizes the asymptotic limit of the inverse of the sum of two different sample covariance matrices.
To motivate our study, consider a special case where $A_1=A_2=1$. 
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
By Proposition 1 of \citet{WZR20}, we consider the case of $r = 1 < t = 2$, since when $r > 1$, the optimal solution of $f(A, B)$ is equivalent to single-task learning.
When $r = 1$, $B$ is a vector and $A_1, A_2$ are both scalars.
Hence we can write down a closed form equation for any local minimum of $f(A, B)$.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align}\label{def hatsig}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \ \ \text{ where } \ \ \hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}.
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariances, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.

To derive the bias and variance equation, we consider the expected loss conditional on the covariates as follows (the empirical loss is close to its expectation similar to Claim \ref{claim_pred_err}):
\begin{align}
	\exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}} =
	& \bignorm{{\Sigma^{(2)}}^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}
Our main result in this section characterizes the asymptotic limit of the above equation under the high-dimensional assumption.
Intuitively, the spectrum of $\hat{\Sigma}^{-1}$ (and hence its trace) not only depends on both tasks' sample sizes, but also depends on the ``alignment'' between $\Sigma^{(1)}$ and $\Sigma^{(2)}$.
However, capturing this intuition quantitatively turns out to be technically challenging.
%The main technical challenge of our result deals with the ``covariate shift'' between tasks one and two.
We introduce a key quantity $M \define (\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$, and as we show below, the trace of $\hat{\Sigma}^{-1}$ has an intricate dependence on the spectrum of $M$.

@@

\begin{corollary}\label{cor_MTL_loss}
	In the setting of Example \ref{ex_sample_ratio}, assume that
	%a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that
	(i) both tasks sample sizes are at least $\frac{3}{2}p$;
	(ii) noise variance is smaller than the shared signal variance: $\sigma^2 \lesssim  \kappa^2$;
%	\be\label{choiceofpara0}
%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
	%\be\label{choiceofpara0}
%	(ii) the task-specific variance of $\beta_i$ is much smaller than the signal strength {\color{red}$d^2 = \oo( {\kappa^2})$}; \HZ{what does $\ll$ mean exactly?}
%	(iii) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$.
	(iii) task-specific variance is much smaller than the shared signal variance: $d^2 \le p^{-\e}{\kappa^2}$.
	Let $\varepsilon = (1 + \sqrt{p/n_1})^ 4 - 1$, which decreases as $n_1$ increases.
	The prediction loss of hard parameter sharing for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le
	  \left|L(\hat{\beta}_2^{\MTL}) - \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} -\frac{\sigma^2 p}{n_1 + n_2 - p}  \right| \le \varepsilon \cdot \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} +  \OO(p^{-c/2}).
	%\left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	%& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + %\frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
	 \end{align}
%	 with high probability for any fixed $c\in(0, \min(\frac{1}{4}, \delta,\frac{\varphi-4}{2\varphi}))$.
%	 {\color{red}[FY: the error also contains $p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) $, both of which cannot be omitted, because (i) there is no assumption on the upper bound of $\kappa^2$, and (ii) we do not necessarily have $c_\varphi<1/4$. We can decide how to present the result concisely (for instance we can impose an upper bound on $\kappa^2$ and that $c_\varphi<1/4$), but it needs to be correct.]}
	 \end{corollary}

>>>>

\paragraph{Bias-variance equations.}
Our key result characterizes the asymptotic limit of the inverse of the sum of two different sample covariance matrices.
To motivate our study, consider a special case where $A_1=A_2=1$. 
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
By Proposition 1 of \citet{WZR20}, we consider the case of $r = 1 < t = 2$, since when $r > 1$, the optimal solution of $f(A, B)$ is equivalent to single-task learning.
When $r = 1$, $B$ is a vector and $A_1, A_2$ are both scalars.
Hence we can write down a closed form equation for any local minimum of $f(A, B)$.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align}\label{def hatsig}
	\hat{\beta}_2^{\MTL} = {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \ \ \text{ where } \ \ \hat{\Sigma} = {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}.
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariances, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.

To derive the bias and variance equation, we consider the expected loss conditional on the covariates as follows (the empirical loss is close to its expectation similar to Claim \ref{claim_pred_err}):
\begin{align}
	\exarg{\cE}{L(\hat{\beta}_2^{\MTL}) \mid X^{(1)}, X^{(2)}} =
	& \bignorm{{\Sigma^{(2)}}^{1/2} \hat{\Sigma}^{-1} {X^{(1)}}^{\top} X^{(1)} (\beta^{(1)} - \beta^{(2)})}^2 \label{eq_bias_2task} \\
	& + \sigma^2 \bigtr{\Sigma^{(2)}\hat{\Sigma}^{-1}}. \label{eq_variance_2task}
\end{align}
Equation \eqref{eq_bias_2task} and \eqref{eq_variance_2task} correspond to the bias and variance of hard parameter sharing for two tasks, respectively.
Our main result in this section characterizes the asymptotic bias-variance limits under the high-dimensional assumption.
Intuitively, the spectrum of $\hat{\Sigma}^{-1}$ (and hence its trace) not only depends on both tasks' sample sizes, but also depends on the ``alignment'' between $\Sigma^{(1)}$ and $\Sigma^{(2)}$.
However, capturing this intuition quantitatively turns out to be technically challenging.
%The main technical challenge of our result deals with the ``covariate shift'' between tasks one and two.
We introduce a key quantity $M \define (\Sigma^{(1)})^{1/2}(\Sigma^{(2)})^{-1/2}$, and as we show below, the trace of $\hat{\Sigma}^{-1}$ has an intricate dependence on the spectrum of $M$.

@@

\begin{corollary}\label{cor_MTL_loss}
	In the setting of Example \ref{ex_sample_ratio}, assume that
	%a) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$, b) $\Sigma_1=\Sigma_2=\id_p$, and c) %there exists a small constant $c_0>0$ such that
	(i) both tasks sample sizes are at least $3p$;
	(ii) noise variance is smaller than the shared signal variance: $\sigma^2 \lesssim  \kappa^2$;
%	\be\label{choiceofpara0}
%	p^{-1/2+c_0}\sigma^2 + p^{c_0}d^2\le \kappa^2\le p^{1-c_0} (\sigma^2 +d^2)  .  	\ee
	%\be\label{choiceofpara0}
%	(ii) the task-specific variance of $\beta_i$ is much smaller than the signal strength {\color{red}$d^2 = \oo( {\kappa^2})$}; \HZ{what does $\ll$ mean exactly?}
%	(iii) the sample sizes $n_1$ and $n_2$ are greater than $(1 + \tau) p$.
	(iii) task-specific variance is much smaller than the shared signal variance: $d^2 \le p^{-\e}{\kappa^2}$.
	Let $\varepsilon = (1 + \sqrt{p/n_1})^ 4 - 1$, which decreases as $n_1$ increases.
	The prediction loss of hard parameter sharing for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le
	  \left|L(\hat{\beta}_2^{\MTL}) - \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} -\frac{\sigma^2 p}{n_1 + n_2 - p}  \right| \le \varepsilon \cdot \frac{d^2 n_1^2 (n_1 + n_2)}{(n_1 + n_2 - p)^3} +  \OO(p^{-c/2}).
	%\left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] d^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} \\
	%& +C \left[(p^{-c_\varphi}+p^{-c_\infty/2})(\sigma^2 +d^2)+p^{-c_\infty}\kappa^2 + %\frac{d^4+\sigma^2 d^2}{\kappa^2}\right],\nonumber
	 \end{align}
%	 with high probability for any fixed $c\in(0, \min(\frac{1}{4}, \delta,\frac{\varphi-4}{2\varphi}))$.
%	 {\color{red}[FY: the error also contains $p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) $, both of which cannot be omitted, because (i) there is no assumption on the upper bound of $\kappa^2$, and (ii) we do not necessarily have $c_\varphi<1/4$. We can decide how to present the result concisely (for instance we can impose an upper bound on $\kappa^2$ and that $c_\varphi<1/4$), but it needs to be correct.]}
	 \end{corollary}
